---
title: "Data Preprocessing for Rat Inspection Dataset"
author: "Jia Ji, Yijia Jiang, Gonghao Liu, Yifei Xu, Ziyan Xu"
date: "2022-11-21"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(RSocrata)
library(rnoaa)
```


First, read in the rat inspection dataset.
```{r, message=FALSE, warning=FALSE}
url_rat = "https://data.cityofnewyork.us/OData.svc/p937-wjvj"
rat = read.socrata(url_rat) %>% 
  janitor::clean_names() 
```


Then we clean the date and keep the records between 2013 and 2021. All missing data are dropped.
```{r, message=FALSE, warning=FALSE}
rat_tidy = rat %>%
  select(inspection_type, bbl, zip_code, latitude, longitude, borough, result, inspection_date, approved_date) %>%
  drop_na() %>% 
  mutate(boro_code = substr(bbl, 1, 1),
         block = substr(bbl, 2, 6),
         lot = substr(bbl, 7, 10)) %>% 
  select(inspection_type, boro_code, block, lot, zip_code, latitude, longitude, borough, result, inspection_date, approved_date) %>% 
  separate(inspection_date, c("inspection_date", "inspection_time"), " ") %>% 
  separate(inspection_date, c("inspection_year", "inspection_month", "inspection_day"), "-") %>%
  separate(approved_date, c("approved_date", "approved_time"), " ") %>% 
  separate(approved_date, c("approved_year", "approved_month", "approved_day"), "-") %>%
  mutate(inspection_year = as.integer(inspection_year), 
         inspection_month = as.integer(inspection_month), 
         inspection_day = as.integer(inspection_day)) %>%
  mutate(approved_year = as.integer(approved_year), 
         approved_month = as.integer(approved_month), 
         approved_day = as.integer(approved_day)) %>%
  relocate(inspection_year, .before = "inspection_month") %>% 
  relocate(approved_year, .before = "approved_month") %>%
  arrange(inspection_year, inspection_month) %>% 
  mutate(inspection_month = month.abb[inspection_month],
         approved_month = month.abb[approved_month]) %>% 
  filter(inspection_year >= 2012 & inspection_year <= 2021)
```


In the next step, we import the weather dataset from NOAA package.
```{r, message=FALSE, warning=FALSE}
# Get the Central Park Station id
nycstationsid = ghcnd_stations() %>% 
  filter(id == "USW00094728") %>% 
  distinct(id)

# Pull the desired weather data for Central Park Station
nyc_weather = meteo_pull_monitors(nycstationsid$id, 
                             date_min = "2012-01-01", 
                             date_max = "2021-12-31",
                             var = c("PRCP", "SNOW", "SNWD", "TMAX", "TMIN"))

nyc_weather_tidy = nyc_weather %>% 
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year),
         month = month.abb[as.numeric(month)],
         day = as.numeric(day)) %>%
  mutate(prcp = prcp/10,
         tmax = tmax/10,
         tmin = tmin/10) 
```


Merge the rat info with weather info.
```{r, message=FALSE, warning=FALSE}
rat_2012_weather = rat_tidy %>% 
  filter(inspection_year == 2012) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id)

rat_2013_weather = rat_tidy %>% 
  filter(inspection_year == 2013) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id)

rat_2014_weather = rat_tidy %>% 
  filter(inspection_year == 2014) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id)

rat_2015_weather = rat_tidy %>% 
  filter(inspection_year == 2015) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id)

rat_2016_weather = rat_tidy %>% 
  filter(inspection_year == 2016) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id)

rat_2017_weather = rat_tidy %>% 
  filter(inspection_year == 2017) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id)

rat_2018_weather = rat_tidy %>% 
  filter(inspection_year == 2018) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id)

rat_2019_weather = rat_tidy %>% 
  filter(inspection_year == 2019) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id)

rat_2020_weather = rat_tidy %>% 
  filter(inspection_year == 2020) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id)

rat_2021_weather = rat_tidy %>% 
  filter(inspection_year == 2021) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id)
```


Thirdly, we import the covid dataset.
```{r, message=FALSE, warning=FALSE}
url_covid = "https://data.cityofnewyork.us/OData.svc/rc75-m7u3"
covid = read.socrata(url_covid) %>% 
  janitor::clean_names() 

covid_tidy = covid %>%
  rename(date = date_of_interest) %>% 
  select(date, contains("case_count")) %>% 
  select(-contains(c("probable_case_count", "case_count_7day_avg", "all_case_count_7day_avg"))) %>%
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year),
         month = month.abb[as.numeric(month)],
         day = as.numeric(day)) %>%
  pivot_longer(
    cols = bx_case_count:si_case_count,
    names_to = "borough",
    values_to = "borough_case_count"
  ) %>% 
  mutate(borough = gsub("_case_count", "", borough)) %>% 
  mutate(borough = recode(borough, "bx" = "Bronx","bk" = "Brooklyn","mn" = "Manhattan","si" = "Staten Island","qn" = "Queens")) %>% 
  relocate(case_count, .after = borough_case_count) %>% 
  rename(total_case_count = case_count) 
```


Merge the above dataframe with covid info.
```{r, message=FALSE, warning=FALSE}
rat_weather_covid = rat_tidy %>% 
  filter(inspection_year == 2020 || 2021) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id) %>% 
  merge(covid_tidy, by.x = c("inspection_year","inspection_month","inspection_day","borough"), by.y = c("year","month","day","borough"))
```


Since the raw dataset is too large, we will export the tidied dataset which is merged with weather info and covid info for further use.
```{r, message=FALSE, warning=FALSE}
write.csv(rat_2012_weather,"rat_2012.csv")
write.csv(rat_2013_weather,"rat_2013.csv")
write.csv(rat_2014_weather,"rat_2014.csv")
write.csv(rat_2015_weather,"rat_2015.csv")
write.csv(rat_2016_weather,"rat_2016.csv")
write.csv(rat_2017_weather,"rat_2017.csv")
write.csv(rat_2018_weather,"rat_2018.csv")
write.csv(rat_2019_weather,"rat_2019.csv")
write.csv(rat_2020_weather,"rat_2020.csv")
write.csv(rat_2021_weather,"rat_2021.csv")
write.csv(rat_weather_covid,"rat_weather_covid.csv")
```

